Paper 1

### 2015-10-17 Articles/Article_EcolEcon_2015/Paper1/Paper1-lambda-alpha.Rnw
Paper1 files.

% This article has been prepared for publication in RStudio with knitr.
% According to http://www.elsevier.com/author-schemas/the-elsarticle-latex-document-class, we should be using the
% elsarticle.cls file.
% According to http://cdn.elsevier.com/assets/pdf_file/0006/109392/journal_refstyles.pdf, we should be using
% elsarticle-template-2-harv.tex as the template for the text.
% Furthermore, we should be using model2-names.bst for the bibliographic references.
% The approach here is to load the frontmatter and backmatter from elsarticle-template-2-harv.tex
% both ahead of and behind the text for our paper.
% -- Matthew Kuperus Heun, 2014-08-01

%% This is file `elsarticle-template-2-harv.tex',
%%
%% Copyright 2009 Elsevier Ltd
%%
%% This file is part of the 'Elsarticle  Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%
%% Template article for Elsevier's document class `elsarticle'
%% with harvard style bibliographic references
%%
%% $Id: elsarticle-template-2-harv.tex 155 2009-10-08 05:35:05Z rishi $
%% $URL: http://lenova.river-valley.com/svn/elsbst/trunk/elsarticle-template-2-harv.tex $
%%
\documentclass[preprint,authoryear,12pt]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

@@

%% if you use PostScript figures in your article
%% use the graphics package for simple commands
%% \usepackage{graphics}
%% or use the graphicx package for more complicated commands
%% \usepackage{graphicx}
%% or use the epsfig package if you prefer to use the old commands
%% \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

\usepackage{wrapfig}    % Allows wrapping of text around figures
\usepackage{soul}       % Provides strikethrough text
\usepackage{float}      % Allows precise positioning of tables and figures within the text
\usepackage{amsmath}    % Allows \begin{equation*} \end{equation*} for unnumbered equations
\usepackage{multirow}   % To create multirow tables
\usepackage{hyperref}   % To create hyperlinks in the paper
\usepackage{microtype}  % produces hanging punctuation and beautiful type
\usepackage{booktabs}   % for beautiful tables
\usepackage{dcolumn}    % aligns table columns at decimal places

% From http://economics.utoronto.ca/osborne/latex/BIBTEX.HTM
\newcommand{\citeapos}[1]{\citeauthor{#1}'s (\citeyear{#1})} % Posessive citations.

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

@@

%%   round  -  round parentheses are used (default)
%%   square -  square brackets are used   [option]
%%   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option>
%%   semicolon  -  multiple citations separated by semi-colon (default)
%%   colon  - same as semicolon, an earlier confusion
%%   comma  -  separated by comma
%%   authoryear - selects author-year citations (default)
%%   numbers-  selects numerical citations
%%   super  -  numerical citations as superscripts
%%   sort   -  sorts multiple citations according to order in ref. list
%%   sort&compress   -  like sort, but also compresses numerical citations
%%   compress - compresses without sorting
%%   longnamesfirst  -  makes first citation full author list
%%
%% \biboptions{longnamesfirst,comma}

@@

<<setup, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE>>=
library(EconData)
library(EconModels)
library(dplyr)
library(reshape2)
library(ggplot2)
library(mosaic)
library(xtable)
library(knitr)
library(texreg)
library(tidyr)
library(scales)
library(tidyr)

# Tell whether to re-run the analysis to create the models
# or the resampling on the (kl)e models.
# This file will use previously-saved models and resampling to generate
# the paper if set to FALSE, greatly reducing time.
# If something changes in the data or in the code that
# performs the analysis or resampling, set 
# rerunModels = TRUE
# rerunResampling = TRUE
# once to re-generate the information needed to 
# create the paper.
# ********************************************************
rerunModels=FALSE
rerunResampling=FALSE
nResamples <- 100 # Number of resamples
# ********************************************************
opts_chunk$set(
  # ********************************************************
  eval=TRUE,
  # ********************************************************
  # tikz allows LaTeX code in graphical output. 
  # E.g., "$y$" for a variable name in the legend for a graph.  
  # dev='tikz',
  tidy=FALSE,
  comment=NA,
  # Tells whether to cache output from chunks are saved. Cacheing saves time. 
  # However, references to LaTeX tables DO NOT WORK if cacheing is turned on.
  cache=FALSE,
  warning=FALSE,  #Tells whether to show warnings in the output.
  # Turns off messages for all chunks. 
  # Set TRUE on an individual chunk to see it.
  message=FALSE, 
  # Tells whether to echo code for all chunks. 
  # Set TRUE on an individual chunk to see its code.
  echo=FALSE,
  # Sets all table captions to the top.
  xtable.caption.placement="top"
)

# This function is used in many places below.
### That's nice.  I assume that's why you wrote a function.  But what does it DO?
@

<<myXTable>>=
myXTable <- function(data, caption, label, digits=2){
  table <- xtable(data, caption=caption, label=label, digits=digits)
  print(table, 
        caption.placement="top", 
        size="\\tiny", 
        include.rownames=FALSE, 
        math.style.negative=TRUE, 
        sanitize.text.function=function(x){x},
        booktabs=TRUE)
}
@

<<model_points_style>>=
model_points_style <- function(p) {
  p + 
  scale_colour_manual(values = c("gray70", "black")) +
  scale_size_manual(values = c(5, 3,3,3,3,3,3)) +
  # scale_alpha_manual(values = c(.2, .6,.6,.6,.6, .6, .6)) +
  scale_fill_manual(values = c(scales::alpha("gray50", .2), rep("transparent", 6))) +
  scale_shape_manual(values = c(21, 21, 23:27)) +
  guides( 
    col = guide_legend(override.aes = aes(size = 5, shape = 21)) )
}

model_points_style_quad <- function(p) {
  p + 
  # scale_colour_manual(values = c("gray70", "black")) +
  scale_size_manual(values = c(5, 3,3,3,3,3,3)) +
  # scale_alpha_manual(values = c(.2, .6,.6,.6,.6, .6, .6)) +
  scale_fill_manual(values = c(scales::alpha("gray50", .2), rep("transparent", 6))) +
  scale_shape_manual(values = c(21, 21, 23:27)) +
  theme(legend.title=element_blank())
}
@

@@

\title{Energy and the CES production function}

@@

\author[IST]{Jo\~{a}o dos Santos}
\author[CalvinEngr]{Matthew Kuperus Heun\corref{cor1}}
\author[Leeds]{Paul Brockway}
\ead{mkh2@calvin.edu}
\author[Leeds]{Marco Sekai}
\author[IST]{Tiago Domingos}
\author[CalvinMath]{Randall Pruim}

\address[IST]{IST, Lisbon, Portugal}
\address[Leeds]{Leeds University, Leeds, UK}
\address[CalvinEngr]{Engineering Department, Calvin College, Grand Rapids, MI 49546, USA}
\address[CalvinMath]{Mathematics \& Statistics Department, Calvin College, Grand Rapids, MI 49546, USA}
\cortext[cor1]{Corresponding author}

\begin{abstract}
%% Text of abstract
**** Add abstract ****
\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
economic growth \sep exergy \sep energy \sep CES
%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)
\end{keyword}

@@

% \linenumbers
%% main text

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction} 
\label{sec:Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This is a study of the effects of quality-adjusted production function inputs
on the importance of energy in the production function.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Coordinates of Analysis} 
\label{sec:Coordinates}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This section describes the coordinates of analysis and briefly reviews 
literature related to each.

%++++++++++++++++++++++++++++++
\subsection{Mathematical Forms of the Energy-augmented Production Function} 
\label{sec:Prod_Func_Forms}
%++++++++++++++++++++++++++++++

In this paper, we assess a prominent energy-augmented production function:
Constant Elasticity of Substitution (CES).
The CES model is assessed relative to 
a model of exponential growth only.
The following subsections describe each.

%------------------------------
\subsubsection{Constant Elasticity of Substitution Production Function (CES)} 
\label{sec:CES}
%------------------------------

The energy-augmented 
Constant Elasticity of Substitution (CES) production function%
  \footnote{
  We use the The \texttt{R} 
  % \citep{R} 
  package \texttt{micEconCES} 
  % \citep{Henningsen:2011td} 
  to estimate the CES production function.
  }
can be given as
%
\begin{equation} \label{eq:CESgeneric}
  y = \gamma \: A \: \left\{\delta \left[\delta_1 x_1^{-\rho_1} 
      + (1-\delta_1)x_2^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) x_3^{-\rho} \right\}^{-1/\rho}; 
      A \equiv \mathrm{e}^{\lambda (t-t_0)},
\end{equation}
%
where
$y \equiv Y/Y_{0}$,
$\gamma$ is a scale parameter,
e is the base of the natural logarithm, 
$\lambda$ is the Solow residual (often thougth to represent the pace of technological progress),
$t$ (time) is measured in years,
$x_1$, $x_2$, and $x_3$ are factors of production
and permutations of capital~($k$), labor~($l$), and energy~($e$),
$k \equiv K/K_{0}$, 
$l \equiv L/L_{0}$, 
$Y$ (economic output) is represented by GDP, 
$K$ (capital) is expressed in currency units, 
$L$ (labor) is expressed in workers or work-hours/year, and
the 0 subscript indicates values at an initial year.%
  \footnote{Dimensionless, indexed quantities are represented by 
  lower-case symbols 
  ($y$, $k$, $l$, $e$, $q$, $x$, and $u$), and dimensional 
  quantities are represented by upper-case symbols 
  ($Y$, $K$, $L$, $E$, $Q$, $X$, and $U$). 
  Model parameters are represented by Greek letters
  ($\alpha_1$, $\alpha_2$, $\lambda$, $\theta$).
  }
The CES model without energy is given in Equation~\ref{eq:CESkl}.
%
\begin{equation} \label{eq:CESkl}
  y = \gamma \: A \: \left[\delta_1 k^{-\rho_1} 
      + (1-\delta_1)l^{-\rho_1} \right]^{-1/\rho_1}; 
      A \equiv \mathrm{e}^{\lambda (t-t_0)} . 
\end{equation}
%
Equation~\ref{eq:CESkle} augments Equation~\ref{eq:CESkl} with energy 
using a ($kl$)($e$) nesting structure, as is typical in the literature. 
Equation~\ref{eq:CESkl} is a degenerate form of Equation~\ref{eq:CESkle} 
where $\delta \rightarrow 1$. 
%
\begin{equation} \label{eq:CESkle}
  y = \gamma \: A \: \left\{\delta \left[\delta_1 k^{-\rho_1} 
      + (1-\delta_1)l^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) e^{-\rho} \right\}^{-1/\rho}; 
      A \equiv \mathrm{e}^{\lambda (t-t_0)}.
\end{equation}
%
In the CES production function, 
$\gamma$ is a fitting parameter that accounts for an atypical first year. 
The fitting parameters $\rho_1$ and $\rho$ indicate elasticities 
of substitution ($\sigma_1$ and $\sigma$).
The elasticity of substitution 
between capital ($k$) and labor ($l$) is given by 
$\sigma_1 = \frac{1}{1+\rho_1}$, and
the elasticity of substitution between ($kl$) and ($e$) is given by 
$\sigma = \frac{1}{1+\rho}$. 
As $\rho_1 \rightarrow 0$, $\sigma_1 \rightarrow 1$,
and the embedded CES production function for $k$ and $l$ degenerates 
to the Cobb-Douglas production function.
Similarly, as $\rho \rightarrow 0$, $\sigma \rightarrow 1$,
and the CES production function for $(kl)$ and $(e)$ degenerates 
to the Cobb-Douglas production function.
As $\sigma \rightarrow \infty$ ($\rho \rightarrow -1$), 
($kl$) and ($e$) are perfect substitutes. 
As $\sigma \rightarrow 0$ ($\rho \rightarrow \infty$), 
($kl$) and ($e$) are perfect complements: 
no substitution is possible. Similarly, 
as $\sigma_1 \rightarrow 0$ ($\rho_1 \rightarrow \infty$),
$k$ and $l$ are perfect complements.
$\delta_1$ describes the relative importance of capital ($k$)
and labor ($l$), and
$\delta$ describes the importance of ($kl$) relative to ($e$).

Constraints on the fitting parameters include 
$\delta_1 \in [0,1]$,
$\delta \in [0,1]$,
$\rho_1 \in [-1,0) \cup (0,\infty)$, and
$\rho \in [-1,0) \cup (0,\infty)$.

Two other nestings of the factors of production ($k$, $l$, and $e$)
are possible with the CES model.
%
\begin{equation} \label{eq:CESlek}
  y = \gamma \: A \: \left\{\delta \left[\delta_1 l^{-\rho_1} 
      + (1-\delta_1) e^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) k^{-\rho} \right\}^{-1/\rho}; 
      A \equiv \mathrm{e}^{\lambda (t-t_0)}
\end{equation}
%
\begin{equation} \label{eq:CESekl}
  y = \gamma \: A \: \left\{\delta \left[\delta_1 e^{-\rho_1} 
      + (1-\delta_1) k^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) l^{-\rho} \right\}^{-1/\rho}; 
      A \equiv \mathrm{e}^{\lambda (t-t_0)}
\end{equation}
% 
Note that $\rho$ ($\sigma$), $\rho_1$ ($\sigma_1$), $\delta$, and $\delta_1$ have different
meanings depending upon the nesting of the factors of production.

For the CES model, factors shares~($\alpha_1$, $\alpha_2$, and $\alpha_3$) are calculated by
$\alpha_1 = \delta_1$, $\alpha_2 = 1 - \delta_1$, and $\alpha_3 = 0$ 
for the CES model with two factors of production
and by 
$\alpha_1 = \delta \delta_1$, $\alpha_2 = \delta(1 - \delta_1)$, and $\alpha_3 = 1 - \delta$
for the CES model with three factors of production and the ($kl$)$e$ nesting.
The values of $\alpha_1$, $\alpha_2$, and $\alpha_3$ are interpreted as
factor shares for capital, labor, and energy, respectively,
as shown in Table \ref{tab:CES_abg_equations}.

\begin{table} \caption{Equations for $\alpha_1$, $\alpha_2$, and $\alpha_3$ 
                        (factor shares for capital, labor, and energy, respectively)
                        for the various CES nestings.} 
\label{tab:CES_abg_equations} 
  \begin{center}
    \begin{tabular}{c c c c} 
      \toprule
      Nesting    & $\alpha_1$              & $\alpha_2$              & $\alpha_3$       \\
      \midrule
      ($kl$) \;  & $\delta_1$              & $1 - \delta_1$          & 0          \\
      ($kl$)$e$  & $\delta \delta_1$       & $\delta (1 - \delta_1)$ & $1 - \delta$ \\   
      ($le$)$k$  & $1 - \delta$            & $\delta \delta_1$       & $\delta (1 - \delta_1)$ \\
      ($ek$)$l$  & $\delta (1 - \delta_1)$ & $1 - \delta$            & $\delta \delta_1$ \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

%------------------------------
\subsubsection{Exponential Production Function (reference model)} 
\label{sec:exp}
%------------------------------

We define an exponential-only reference model for economic growth
in Equation~\ref{eq:exponential-only-model}.
%
\begin{equation} \label{eq:exponential-only-model}
  y = \theta A \; ; \; A \equiv \mathrm{e}^{\lambda(t-t_0)} \; .
\end{equation}
%
We expect that the reference model
will have a larger fitted Solow residual term than
the CES model, because no factors of production are
included in the reference model to drive economic growth.
Indeed, in the reference model, 
all economic growth is attributed to the Solow residual~($\lambda$).

In contrast, it is not necessarily true that 
CES models will exhibit lower $sse$ than the reference model
shown in Equation~\ref{eq:exponential-only-model}.
The CES model has more fitting parameters,
but it incorporates the factors of production
at constant returns to scale. 
The reference model has $\alpha_1 + \alpha_2 + \alpha_3 = 0$,
whereas the CES model~(Equation~\ref{eq:CESgeneric})
implicitly assumes constant returns to scale.
Thus, if the factors of production~($\alpha_1$, $\alpha_2$, and $\alpha_3$) 
are poorly correlated to output~($y$),
the constant returns to scale constraint may cause higher $sse$ for
the CES model relative to the reference model.

%++++++++++++++++++++++++++++++
\subsection{Economies} 
\label{sec:Economies}
%++++++++++++++++++++++++++++++

Discuss economies here.

UK and Portugal.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Sources of Data}
\label{sec:Sources_of_Data}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Discuss data sources here.

%++++++++++++++++++++++++++++++
\subsection{Historical Data} 
\label{sec:Historical_Data}
%++++++++++++++++++++++++++++++

Historical data are stored in the \texttt{IST} and \texttt{Leeds} data sets.

<<Formulas>>=
# Set up the formulas for each type of analysis.
# These formulas contain energy. 
# We can use keep_summands to drop the energy term (3rd summand on the rhs)
# later if needed.
formulas <- 
  list( 
    PT = c(unadjusted = build_formula("iGDP", list("iKstkS.L", "iL", "iXpMP", "iYear")),
             adjusted = build_formula("iGDP", list("iKservS.L", "ihLest", "iUMP", "iYear"))),
    UK = c(unadjusted = build_formula("iGDP", list("iKstkO.WwithRD", "iL", "iXp", "iYear")),
             adjusted = build_formula("iGDP", list("iKservO.WwithRD", "ihLest", "iU", "iYear")))
  )
countries <- names(formulas)
flavors <- names(formulas$PT)
@

<<Data>>=
# Create the historicalData data frames
### If we need to select variables, we can do that.  Unfortunately, select() and select_() don't do
### what we would like here (they don't accept a vector of character strings), so we'd have to use
### something else.  I've kept the fake select_() command here just to record what was done earlier.

commonVars <- c("Year", "Country", "Source")
historicalData <- list(
  PT =  IST %>% 
    filter(Country == "PT"), 
    #  %>% select_(unique(c(commonVars, all.vars(formulas$PT$unadjusted), all.vars(formulas$PT$adjusted)))),
  UK = Leeds %>% 
    filter(Country == "UK") 
    # %>% select_(unique(c(commonVars, all.vars(formulas$UK$unadjusted), all.vars(formulas$UK$adjusted))))
)
@

For unadjusted variables, we use the time series shown in Table~\ref{tab:unadjusted_variables}.

\begin{table} \caption{Unadjusted time-series variables.} 
\label{tab:unadjusted_variables} 
  \begin{center}
    \begin{tabular}{r l l} 
      \toprule
      Variable    & Portugal           & UK                 \\
      \midrule
      Output      & \Sexpr{summands(formulas$PT$unadjusted, n=1, right=FALSE)}   
                                                             & \Sexpr{summands(formulas$UK$unadjusted, n=1, right=FALSE)}\\
      Capital     & \Sexpr{summands(formulas$PT$unadjusted, 1)} & \Sexpr{summands(formulas$UK$unadjusted, 1)} \\
      Labor       & \Sexpr{summands(formulas$PT$unadjusted, 2)} & \Sexpr{summands(formulas$UK$unadjusted, 2)} \\
      Energy      & \Sexpr{summands(formulas$PT$unadjusted, 3)} & \Sexpr{summands(formulas$UK$unadjusted, 3)} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

For quality-adjusted variables, we use the time series shown in Table~\ref{tab:quality_adjusted_variables}.

\begin{table} \caption{Quality-adjusted time series variables.} 
\label{tab:quality_adjusted_variables} 
  \begin{center}
    \begin{tabular}{r l l} 
      \toprule
      Variable    & Portugal         & UK               \\
      \midrule
      Output      & \Sexpr{summands(formulas$PT$adjusted, n=1, right=FALSE)}   
                                                           & \Sexpr{summands(formulas$UK$adjusted, n=1, right=FALSE)} \\
      Capital     & \Sexpr{summands(formulas$PT$adjusted, 1)} & \Sexpr{summands(formulas$UK$adjusted, 1)} \\
      Labor       & \Sexpr{summands(formulas$PT$adjusted, 2)} & \Sexpr{summands(formulas$UK$adjusted, 2)} \\
      Energy      & \Sexpr{summands(formulas$PT$adjusted, 3)} & \Sexpr{summands(formulas$UK$adjusted, 3)} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Parameter Estimation (Without Cost-share Theorem)}
\label{sec:parameter_estimation_noCSP}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Parameter estimation is accomplished with \texttt{formula}s of the form
% %
% \begin{equation} \label{eq:formulas}
%   y \sim x_1 + x_2 + x_3 + time,
% \end{equation}
% %
% where $y$ is economic output and 
% $x_1$, $x_2$, and $x_3$ are factors of production
% and permutations of
% capital ($k$), labor ($l$), and energy ($e$).
% (The factors of production may be quality-adjusted.)
% $time$ is the time variable, usually \texttt{iYear}. 
% The \texttt{formula}s are used in a fitting function such as 
% \texttt{cdModel} to create a \texttt{model} object.
% The function \texttt{naturalCoef} extracts reasonable
% coefficients from \texttt{model} objects.
% 
% Note that in the fitting process for the Cobb-Douglas equation,
% $\texttt{scale} = \theta$, 
% $\alpha_1 = \alpha$,
% $\alpha_2 = \beta$, and
% $\alpha_3 = \gamma$,
% when the \texttt{formula} is specified as 
% %
% \begin{equation} \label{eq:formula_kle}
%   \texttt{y} \sim \texttt{k + l + e + time}.
% \end{equation}
% %
% Note that for the CES model, $\texttt{scale} = \gamma$ and is expected to be close to unity.

The models being fit can be described by the algebraic form of the model 
(Cobb-Douglas, CES, etc.) and a formula that enumerates 
which data variables are being used in which roles to fit the model.  
A formula of the form
%
\begin{equation} \label{eq:formulas-3-factor}
y \sim x_1 + x_2 + x_3 + t,
\end{equation}
%
or 
%
\begin{equation} \label{eq:formulas-2-factor}
y \sim x_1 + x_2 + t,
\end{equation}
%
describes the economic output variable ($y$, usually \texttt{iGDP},
                                        indexed GDP) 
and the factors of production ($x_1$, $x_2$, and $x_3$,
                               which will be some measure of capital, labor, and energy, 
                               but perhaps not in that order), 
and a time variable ($t$, usually \texttt{iYear}, 
                     the number of years since the beginning of data collection).  
All of the models assume an error term that is additive on the 
logarithmic scale and are fit by the method of least squares.  
Model fitting provides estimates for all parameters in the model.

<<Models>>=
fileNameCDCES <- "EcolEcon2015ModelsCDCES.rds"
fileNameCSP <- "EcolEcon2015ModelsCSP.rds"
fileNameExp <- "EcolEcon2015ModelsExp.rds"
if (!rerunModels){
  # load from the file on disk
  models <- readRDS(fileNameCDCES)
  modelsCSP <- readRDS(fileNameCSP)
  modelsExp <- readRDS(fileNameExp)
} else {   ### PLEASE USE COMMENTS TO MARK WHERE LONG CHUNKS END; BETTER: AVOID LONG CHUNKS, WHERE POSSIBLE

  ### This has not been inspected by rjp

  # re-run the models and save to disk.
  # In RStudio, be sure to set the working directory to the source file directory.

  #
  # Models that are exponential only, y = A * exp(lambda*t)
  #
  modelsExp <- list()
  alphaExp <- 0
  betaExp <- 0
  gammaExp <- 0
  for (country in countries){
    # Replace with better code later. 
    # Would like to say cdModel(formula, data, factorShares=c(alpha_1 = 0, alpha_2 = 0, alpha_3 = 0))
    # or something like that.
    form <- keep_summands(formulas[[country]][["unadjusted"]], n = c(-1, -2, -3)) # Keep none of the factors of production
    yVar <- as.name(summands(form, n=1, right=FALSE))
    tVar <- as.name(summands(form, n=1))
    data <- historicalData[[country]]
    expMod <- lm(log(eval(yVar)) ~ eval(tVar), data=data)
    attr(expMod, "naturalCoeffs") <- data.frame(lambda=coef(expMod)[[2]],
                                                  logscale=coef(expMod)[[1]], 
                                                  scale=exp(coef(expMod)[[1]]),
                                                  alpha_1=alphaExp,
                                                  alpha_2=betaExp, 
                                                  alpha_3=gammaExp)
    expMod$formula <- form
    sdata <- subset(data, select=all.vars(form))
    sdata <- data[complete.cases(sdata), unique(c(all.vars(form), names(data)))]
    expMod$data <- sdata
    expMod$response <- sdata$iGDP 
    modelsExp[[country]] <- expMod
  }
  # Save exponential models to disk.
  saveRDS(modelsExp, file=fileNameExp)

  #
  # Models that adhere to the cost-share theorem
  #

  modelsCSP <- list()
  # Output elasticities based on cost share theorem
  alphaCSP <- 0.3 # Fraction of national accounts spent on capital
  betaCSP <- 0.7 # Fraction of national accounts spent on labor
  for (country in countries){
    for (adj in c("unadjusted", "adjusted")){
      # Replace with better code later. 
      # Would like to say cdModel(formula, data, factorShares=c(alpha_1 = 0.3, alpha_2 = 0.7, alpha_3 = 0))
      # or something like that.
      form <- keep_summands(formulas[[country]][[adj]], n=-3) # Remove energy
      yVar <- as.name(summands(form, n=1, right=FALSE))
      kVar <- as.name(summands(form, n=1))
      lVar <- as.name(summands(form, n=2))
      tVar <- as.name(summands(form, n=3))
      data <- historicalData[[country]]
      # Cobb-Douglas
      ### breaking a formerly long line
      cdModCSP <- 
        lm(log(eval(yVar)) - alphaCSP*log(eval(kVar)) - betaCSP*log(eval(lVar)) ~ eval(tVar), 
           data = data)
      attr(cdModCSP, "naturalCoeffs") <- data.frame(lambda=coef(cdModCSP)[[2]],
                                                    logscale=coef(cdModCSP)[[1]], 
                                                    scale=exp(coef(cdModCSP)[[1]]),
                                                    alpha_1=alphaCSP,
                                                    alpha_2=betaCSP, 
                                                    alpha_3=0)
      cdModCSP$formula <- form
      sdata <- subset(data, select=all.vars(form))
      sdata <- data[complete.cases(sdata), unique(c(all.vars(form), names(data)))]
      cdModCSP$data <- sdata
      cdModCSP$response <- sdata$iGDP
      modelsCSP[[country]][[adj]][["noE"]][["CD"]] <- cdModCSP
      # CES
      cesModCSP <- plm(log(eval(yVar) / (alphaCSP*eval(kVar)^(-rho_1) + betaCSP*eval(lVar)^(-rho_1))^(-1/rho_1)) ~ eval(tVar),
                       data = historicalData[[country]],
                       params=c(rho_1=0.8))
      names(cesModCSP$coefficients) <- c("logscale", "lambda", "rho_1")
      cesModCSP$coefficients <- c(coef(cesModCSP), c(delta_1=alphaCSP, delta=1))
      attr(cesModCSP, "naturalCoeffs") <- EconModels:::makeNatCoef(cesModCSP)
      cesModCSP$formula <- form
      cesModCSP$data <- sdata
      modelsCSP[[country]][[adj]][["noE"]][["CES"]][["kl"]] <- cesModCSP
    }
  }
  # Save cost-share theorem models to disk.
  saveRDS(modelsCSP, file=fileNameCSP)

  #
  # Models where all parameters are estimated (models don't adhere to the cost share theorem)
  #

  models <- list()
  for(country in countries){
    for (flavor in flavors){
      fnoE <- keep_summands(formulas[[country]][[flavor]], -3) # Drop the energy term (3rd term on RHS)
      fwithE <- formulas[[country]][[flavor]]
      # CD models
      models[[country]][[flavor]][["noE"]][["CD"]]   <- cdModel(fnoE, data = historicalData[[country]])
      models[[country]][[flavor]][["withE"]][["CD"]] <- cdModel(fwithE, data = historicalData[[country]])
      # CES models
      models[[country]][[flavor]][["noE"]][["CES"]][["kl"]] <- cesModel(fnoE, 
                                                                        data = historicalData[[country]], 
                                                                        nest=c(1,2), digits = 10)
      nestsWithE <- list(kle=c(1,2,3), lek=c(2,3,1), ekl=c(3,1,2))
      models[[country]][[flavor]][["withE"]][["CES"]] <- lapply(nestsWithE, function(nest){
          cesModel(formula=fwithE, data=historicalData[[country]], nest=nest, digits = 10)
        })
    }      
  }
  # Save models to disk.
  saveRDS(models, file=fileNameCDCES)
}  ### END ELSE FROM WAY UP THERE???
@

<<Models2>>=
### TIME FOR A NEW R CHUNK -- That helps detect the end of that long code chunk

### This function is the same in each model group below.  DRY (Don't repeat yourself).
### Using this makes it (a) easier to see what's going on, and (b) easier to make adjustments
### that affect all of the situations simultaneously.

### General hint:  If you find yourself wanting to copy and paste a chunk of code, there is usually
### a better way to do what you are about to do.

>>>>

%% This is file `elsarticle-template-2-harv.tex',
%%
%% Copyright 2009 Elsevier Ltd
%%
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%
%% Template article for Elsevier's document class `elsarticle'
%% with harvard style bibliographic references
%%
%% $Id: elsarticle-template-2-harv.tex 155 2009-10-08 05:35:05Z rishi $
%% $URL: http://lenova.river-valley.com/svn/elsbst/trunk/elsarticle-template-2-harv.tex $
%%
\documentclass[preprint,10pt,3p]{elsarticle}
\usepackage{setspace}
\doublespacing
%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

@@

%% if you use PostScript figures in your article
%% use the graphics package for simple commands
\usepackage{graphics}
%% or use the graphicx package for more complicated commands
\usepackage{graphicx}
%% or use the epsfig package if you prefer to use the old commands
\usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
\bibpunct{(}{)}{,}{a}{,}{,}

\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{amsthm}
\usepackage{float}
\usepackage{epstopdf}
\usepackage{textcomp}
\usepackage{eurosym}
\usepackage{morefloats}
\usepackage{graphicx}
\usepackage{rotating}
\usepackage{caption}
\usepackage{csquotes}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{booktabs}
\usepackage{rotating}

\usepackage{hyperref}

%\usepackage[hidenotes]{authNote}
\usepackage[shownotes]{authNote}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
\usepackage[right]{lineno}

@@

%%   round  -  round parentheses are used (default)
%%   square -  square brackets are used   [option]
%%   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option>
%%   semicolon  -  multiple citations separated by semi-colon (default)
%%   colon  - same as semicolon, an earlier confusion
%%   comma  -  separated by comma
%%   authoryear - selects author-year citations (default)
%%   numbers-  selects numerical citations
%%   super  -  numerical citations as superscripts
%%   sort   -  sorts multiple citations according to order in ref. list
%%   sort&compress   -  like sort, but also compresses numerical citations
%%   compress - compresses without sorting
%%   longnamesfirst  -  makes first citation full author list
%%
%\biboptions{longnamesfirst,comma}

@@

@@

\title{Paper 1}

@@

\author[label1]{Randall J. Pruim}
\ead{rpruim@calvin.edu}

\author[label2]{Matthew K. Heun}
\ead{mkh2@calvin.edu}

\address[label1]{Mathematics \& Statistics Department, Calvin College, Grand
Rapids, MI 49546, USA.}

\address[label2]{Engineering Department, Calvin College, Grand Rapids, MI 49546,
USA.}

\linenumbers

\begin{abstract}

Need abstract here.

\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
CES function \sep Cobb-Douglas function \sep statistical precision \sep aggregate production function
\end{keyword}

@@

\linenumbers

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Highlights}
\label{sec:highlights}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\emph{This will go in a separate document eventually.}

\begin{itemize}
\item Highlight 1
\item Highlight 2
\item Highlight 3
\item Highlight 4
\end{itemize}

%\linenumbers

%%%% Import Analysis chunks

<<child="../EcolEcon_2015A.Rnw", eval=TRUE>>=
@

<<include=FALSE>>=
knitr::opts_chunk$set(include = TRUE)
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Tables}
\label{sec:tables}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<CES_sr_gof, results="asis">>=
data_CES_SR_gof <- 
  table_data %>% 
  filter(Model=="CES" | Model=="Ref.") %>% 
  select(Country, Model, CSP, flavor, Energy, Nest, scale, lambda, sse) %>%
  plyr::rename(c("flavor" = "Factors of prod.",
                  "scale" = "$\\theta$",
                  "lambda" = "$\\lambda$"))
